---
title: "Homework 6"
author: "Ainsel Levitskaia-Collins, HL2710"
date: "2025-11-22"
output: github_document
---

```{r include = FALSE}
library(tidyverse)
library(p8105.datasets)
data("weather_df")
```

### Problem 2

#### Bootstrapping

Function for sampling for bootstrapping:

```{r}
#' Sample for Bootstrapping
#'
#' @param df Dataframe to sample from
#'
#' @returns A sample selected with replacement for bootstrapping
boot_sample = function(df) {
  sample_frac(df, replace = TRUE)
}
```

Tidying and pulling bootstrap samples from `weather_df`:

```{r}
weather_df <- weather_df %>% 
  janitor::clean_names()

weather_bootstrap <- tibble(strap_number = 1:5000) %>% 
  mutate(
    strap_sample = map(strap_number, \(i) boot_sample(df = weather_df))
  )
```

Creating a variable for $\frac{\hat{\beta}_{1}}{\hat{\beta}_{2}}$:

```{r}
weather_bs_betas <- weather_bootstrap %>% 
  mutate(
    models = map(strap_sample, \(df) lm(tmax ~ tmin + prcp, data = df)),
    results = map(models, broom::tidy)
  ) %>% 
  select(-strap_sample, -models) %>% 
  unnest(results) %>% 
  select(strap_number, term, estimate) %>% 
  pivot_wider(
    names_from = term,
    values_from = estimate
  ) %>% 
  mutate(beta_ratio = tmin / prcp)

head(weather_bs_betas)
```

Collecting $\hat{r}^{2}$ data:

```{r}
weather_bs_rsquared <- weather_bootstrap %>% 
  mutate(
    models = map(strap_sample, \(df) lm(tmax ~ tmin + prcp, data = df)),
    results = map(models, broom::glance)
  ) %>% 
  select(-strap_sample, -models) %>% 
  unnest(results)

head(weather_bs_rsquared)
```

Merging the two `weather_bs` datasets, so that all the information is in one place:

```{r}
weather_bs <- merge(weather_bs_betas, weather_bs_rsquared, by = "strap_number") %>% 
  select(strap_number, beta_ratio, r.squared)

head(weather_bs)
```

#### Plot of the ratio of betas vs r-squared

```{r message = FALSE}
plot_weather_estimates <- weather_bs %>% 
  ggplot(aes(x = beta_ratio, y = r.squared)) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE) +
  theme_minimal() +
  xlab("Ratio of beta(minimum temperature) to beta(precipitation)") +
  ylab("R-squared")

plot_weather_estimates
```

As the ratio between the minimum temperature $\hat{\beta}_{1}$ and precipitation $\hat{\beta}_{2}$ decreases, the $\hat{r}^{2}$ value tends to increase.

#### Confidence Interval

```{r}
weather_bs_ci <- tibble(
  beta_ratio = quantile(pull(weather_bs, beta_ratio), probs = c(0.025, 0.975)),
  r_squared = quantile(pull(weather_bs, r.squared), probs = c(0.025, 0.975))
)
```

The 95% confidence interval for $\hat{r}^{2}$ is (`r round(pull(weather_bs_ci, r_squared)[[1]], digits = 2)`, `r round(pull(weather_bs_ci, r_squared)[[2]], digits = 2)`). The 95% confidence interval for $\frac{\hat{\beta}_{1}}{\hat{\beta}_{2}}$ is (`r round(pull(weather_bs_ci, beta_ratio)[[1]], digits = 1)`, `r round(pull(weather_bs_ci, beta_ratio)[[2]], digits = 1)`).



### Problem 3

Loading and cleaning `birthweight` dataset, including renaming variables to something less confusing and lightly reorganizing the variable order:

```{r message = FALSE}
birthweight <- read_csv("./data/birthweight.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    babysex = case_match(babysex,
                         1 ~ "Male",
                         2 ~ "Female"),
    frace = case_match(frace,
                       1 ~ "White",
                       2 ~ "Black",
                       3 ~ "Asian",
                       4 ~ "Puerto Rican",
                       8 ~ "Other",
                       9 ~ "Unknown"),
    mrace = case_match(mrace,
                       1 ~ "White",
                       2 ~ "Black",
                       3 ~ "Asian",
                       4 ~ "Puerto Rican",
                       8 ~ "Other")) %>% 
  rename(
    b_sex = babysex, 
    b_head_cents = bhead, 
    b_length_cents = blength, 
    b_weight_grams = bwt, 
    del_weight_lbs = delwt, 
    fam_income_hunds = fincome, 
    f_race = frace, 
    b_gweeks = gaweeks, 
    b_malform = malform,
    m_menarche_yrs = menarche, 
    m_height_ins = mheight, 
    m_age_yrs = momage, 
    m_race = mrace, 
    num_parity = parity, 
    num_lbw = pnumlbw, 
    num_sga = pnumsga, 
    m_ppbmi = ppbmi, 
    m_ppweight_lbs = ppwt, 
    m_cigs = smoken, 
    m_wtgain_lbs = wtgain) %>% 
  select(b_sex, b_head_cents, b_length_cents, b_weight_grams, b_gweeks, 
         b_malform, del_weight_lbs, fam_income_hunds, f_race, m_menarche_yrs, 
         m_height_ins, m_age_yrs, m_race, m_ppbmi, m_ppweight_lbs, m_cigs, 
         m_wtgain_lbs, num_parity, num_lbw, num_sga)

head(birthweight)
```

#### My model; comparing birth weight against the mother's pre-pregnancy BMI and cigarettes smoked during pregnancy

This model is based on the theory that a mother of poor health (either too low or too high of a BMI, and smoking during pregnancy) will contribute to the potential of a baby's birth weight being unusual.

Fitting the model:

```{r}
birthweight_model1 <- lm(b_weight_grams ~ m_ppbmi + m_cigs, data = birthweight)

birthweight_model1 %>% 
  broom::tidy() %>% 
  mutate(term = case_match(term,
                           "(Intercept)" ~ "Intercept",
                           "m_ppbmi" ~ "Pre-Pregnancy BMI",
                           "m_cigs" ~ "Cigarettes Smoked During Pregnancy")) %>% 
  knitr::kable(digits = 2, col.names = c("Term", "Estimate", "Standard Error", "Statistic", "P-Value"))
```

Testing model fit:

```{r model1, fig.show = "hold", out.width = "50%"}
birthweight_model1_rplot <- birthweight %>% 
  modelr::add_residuals(birthweight_model1) %>% 
  ggplot(aes(x = b_weight_grams, y = resid)) +
  geom_point(alpha = 0.5, shape = 1) +
  theme_minimal() +
  xlab("Birth Weight (grams)") +
  ylab("Residuals")

birthweight_model1_rplot

birthweight_model1_rpplot <- birthweight %>% 
  modelr::add_residuals(birthweight_model1) %>% 
  modelr::add_predictions(birthweight_model1) %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = 0.5, shape = 1) +
  theme_minimal() +
  xlab("Predicted") +
  ylab("Residuals")

birthweight_model1_rpplot
```

This model, when considered independently, looks like a good fit for the data; when comparing the residuals against the birth weight variable, the residuals follow a very linear path with few outliers. When compared against predicted values, the shape of the residuals plot is generally symmetrical and clustered at the center of the plot with no clear pattern.

#### Test Model 1: birthweight compared against length at birth and gestational age (no interaction)

Fitting the model:

```{r}
birthweight_model2 <- lm(b_weight_grams ~ b_length_cents + b_gweeks, data = birthweight)

birthweight_model2 %>% 
  broom::tidy() %>% 
  mutate(term = case_match(term,
                           "(Intercept)" ~ "Intercept",
                           "b_length_cents" ~ "Length at Birth (centimeters)",
                           "b_gweeks" ~ "Gestational Weeks")) %>% 
  knitr::kable(digits = 2, col.names = c("Term", "Estimate", "Standard Error", "Statistic", "P-Value"))
```

Testing the model fit:

```{r model2, fig.show = "hold", out.width = "50%"}
birthweight_model2_rplot <- birthweight %>% 
  modelr::add_residuals(birthweight_model2) %>% 
  ggplot(aes(x = b_weight_grams, y = resid)) +
  geom_point(alpha = 0.5, shape = 1) +
  theme_minimal() +
  xlab("Birth Weight (grams)") +
  ylab("Residuals")

birthweight_model2_rplot

birthweight_model2_rpplot <- birthweight %>% 
  modelr::add_residuals(birthweight_model2) %>% 
  modelr::add_predictions(birthweight_model2) %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = 0.5, shape = 1) +
  theme_minimal() +
  xlab("Predicted") +
  ylab("Residuals")

birthweight_model2_rpplot
```

This model, when considered independently, does not look like a good fit for the data; when comparing the residuals against the birth weight variable, the residuals do not follow a perfect linear path and see some spread. When compared against predicted values, the residuals cluster off-center, with a few notable outliers.